---
title: "Michelson Speed-of-light Measurements"
author: "(Your name here)"
date: 2020-
output: 
  github_document:
    toc: true
prerequisites:
  - e-data02-derive
editor_options: 
  markdown: 
    wrap: 72
---

*Purpose*: When studying physical problems, there is an important
distinction between *error* and *uncertainty*. The primary purpose of
this challenge is to dip our toes into these factors by analyzing a real
dataset.

*Reading*: [Experimental Determination of the Velocity of
Light](https://play.google.com/books/reader?id=343nAAAAMAAJ&hl=en&pg=GBS.PA115)
(Optional)

<!-- include-rubric -->

# Grading Rubric

<!-- -------------------------------------------------- -->

Unlike exercises, **challenges will be graded**. The following rubrics
define how you will be graded, both on an individual and team basis.

## Individual

<!-- ------------------------- -->

| Category | Needs Improvement | Satisfactory |
|----|----|----|
| Effort | Some task **q**'s left unattempted | All task **q**'s attempted |
| Observed | Did not document observations, or observations incorrect | Documented correct observations based on analysis |
| Supported | Some observations not clearly supported by analysis | All observations clearly supported by analysis (table, graph, etc.) |
| Assessed | Observations include claims not supported by the data, or reflect a level of certainty not warranted by the data | Observations are appropriately qualified by the quality & relevance of the data and (in)conclusiveness of the support |
| Specified | Uses the phrase "more data are necessary" without clarification | Any statement that "more data are necessary" specifies which *specific* data are needed to answer what *specific* question |
| Code Styled | Violations of the [style guide](https://style.tidyverse.org/) hinder readability | Code sufficiently close to the [style guide](https://style.tidyverse.org/) |

## Submission

<!-- ------------------------- -->

Make sure to commit both the challenge report (`report.md` file) and
supporting files (`report_files/` folder) when you are done! Then submit
a link to Canvas. **Your Challenge submission is not complete without
all files uploaded to GitHub.**

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(googlesheets4)

url <- "https://docs.google.com/spreadsheets/d/1av_SXn4j0-4Rk0mQFik3LLr-uf0YdA06i3ugE6n-Zdo/edit?usp=sharing"

# Parameters
LIGHTSPEED_VACUUM    <- 299792.458 # Exact speed of light in a vacuum (km / s)
LIGHTSPEED_MICHELSON <- 299944.00  # Michelson's speed estimate (km / s)
LIGHTSPEED_PM        <- 51         # Michelson error estimate (km / s)
```

*Background*: In 1879 Albert Michelson led an experimental campaign to
measure the speed of light. His approach was a development upon the
method of Foucault[3], and resulted in a new estimate of
$v_0 = 299944 \pm 51$ kilometers per second (in a vacuum). This is very
close to the modern *exact* value of `r LIGHTSPEED_VACUUM`. In this
challenge, you will analyze Michelson's original data, and explore some
of the factors associated with his experiment.

I've already copied Michelson's data from his 1880 publication; the code
chunk below will load these data from a public googlesheet.

*Aside*: The speed of light is *exact* (there is **zero error** in the
value `LIGHTSPEED_VACUUM`) because the meter is actually
[*defined*](https://en.wikipedia.org/wiki/Metre#Speed_of_light_definition)
in terms of the speed of light!

```{r read-sheet}
## Note: No need to edit this chunk!
gs4_deauth()
ss <- gs4_get(url)
df_michelson <-
  read_sheet(ss) %>%
  select(Date, Distinctness, Temp, Velocity) %>%
  mutate(Distinctness = as_factor(Distinctness))

df_michelson %>% glimpse()
```

*Data dictionary*:

-   `Date`: Date of measurement
-   `Distinctness`: Distinctness of measured images: 3 = good, 2 = fair,
    1 = poor
-   `Temp`: Ambient temperature (Fahrenheit)
-   `Velocity`: Measured speed of light (km / s)

### **q1** Re-create the following table (from Michelson (1880), pg. 139) using `df_michelson` and `dplyr`. Note that your values *will not* match those of Michelson *exactly*; why might this be?

| Distinctness | n   | MeanVelocity |
|--------------|-----|--------------|
| 3            | 46  | 299860       |
| 2            | 39  | 299860       |
| 1            | 15  | 299810       |

```{r q1-task}
## TODO: Compute summaries
df_q1 <- df_michelson %>% 
  group_by(Distinctness) %>% 
  summarize(
    n = n(),
    MeanVelocity = mean(Velocity)
  )

  
df_q1 %>%
  arrange(desc(Distinctness)) %>%
  knitr::kable()
```

**Observations**:

-   Write your observations here!

    -   The data in the first two columns are identical to the given
        table. The Mean Velocity Values are off by a few numbers.

-   Why might your table differ from Michelson's?

    -   It could be taking different amount into their calculation of
        the mean data. Potentially when transferring the data we forgot
        a few lines. I assume it is due to different ways of calculating
        mean or difference in the data we are analyzing.
    -   This error could also be due to difference in methods of
        rounding with the data. That would explain the why the numbers
        are only slightly off (about 1-2)

The `Velocity` values in the dataset are the speed of light *in air*;
Michelson introduced a couple of adjustments to estimate the speed of
light in a vacuum. In total, he added $+92$ km/s to his mean estimate
for `VelocityVacuum` (from Michelson (1880), pg. 141). While the
following isn't fully rigorous ($+92$ km/s is based on the mean
temperature), we'll simply apply this correction to all the observations
in the dataset.

### **q2** Create a new variable `VelocityVacuum` with the $+92$ km/s adjustment to `Velocity`. Assign this new dataframe to `df_q2`.

```{r q2-task}
## TODO: Adjust the data, assign to df_q2
df_q2 <- df_michelson %>% 
  mutate(
  VelocityVacuum = Velocity + 92
  ) 

df_q2 %>% 
  arrange(desc(Distinctness))
```

As part of his study, Michelson assessed the various potential sources
of error, and provided his best-guess for the error in his
speed-of-light estimate. These values are provided in
`LIGHTSPEED_MICHELSON`---his nominal estimate---and
`LIGHTSPEED_PM`---plus/minus bounds on his estimate. Put differently,
Michelson believed the true value of the speed-of-light probably lay
between `LIGHTSPEED_MICHELSON - LIGHTSPEED_PM` and
`LIGHTSPEED_MICHELSON + LIGHTSPEED_PM`.

Let's introduce some terminology:[2]

-   **Error** is the difference between a true value and an estimate of
    that value; for instance `LIGHTSPEED_VACUUM - LIGHTSPEED_MICHELSON`.
-   **Uncertainty** is an analyst's *assessment* of the error.

Since a "true" value is often not known in practice, one generally does
not know the error. The best they can do is quantify their degree of
uncertainty. We will learn some means of quantifying uncertainty in this
class, but for many real problems uncertainty includes some amount of
human judgment.[2]

### **q3** Compare Michelson's speed of light estimate against the modern speed of light value. Is Michelson's estimate of the error (his uncertainty) greater or less than the true error?

```{r q3-task}
## TODO: Compare Michelson's estimate and error against the true value
## Your code here!

tibble(
  Michelson_LIGHTSPEED = LIGHTSPEED_MICHELSON,
  LowerBound = LIGHTSPEED_MICHELSON - LIGHTSPEED_PM,
  UpperBound = LIGHTSPEED_MICHELSON + LIGHTSPEED_PM,
  Vacuum_LIGHTSPEED = LIGHTSPEED_VACUUM,
  Error = LIGHTSPEED_VACUUM - LIGHTSPEED_MICHELSON,
  Ratio = LIGHTSPEED_PM / Error
)


# print(LowerBound)
# print(UpperBound)
# print(LIGHTSPEED_VACUUM)
```

**Observations**:

-   Is Michelson's estimate of the error (his uncertainty) greater or
    less than the true error?

    -   Michelson's uncertainty is greater than the true error. His
        speed and the lower/upper bound do not include the actual vacuum
        light speed. There is an Error of -151.54.

-   Make a quantitative comparison between Michelson's uncertainty and
    his error.

    -   Michelson's data is uncertain as in all of its measurements has
        a +- 51 value range from the average. In addition to this none
        of the data measurements are in the bounds of the Vacuum Light
        Speed. Michelson's error was larger than his estimation of the
        error (aka his uncertainty). His data was off from the actual
        value but in the grand scheme of scale the numbers are
        surprisingly close.
    -   The ratio between uncertainty and error is 0.336. This means the
        uncertainty is not great enough to account for the full error in
        the measurements. The uncertainty can account for what could be
        considered 33% of the error. aka the uncertainty can take the
        results to be closer to the actual speed of light but does not
        make them the same.

The following plot shows all of Michelson's data as a [control
chart](https://en.wikipedia.org/wiki/Control_chart); this sort of plot
is common in manufacturing, where it is used to help determine if a
manufacturing process is under [statistical
control](https://en.wikipedia.org/wiki/Statistical_process_control).
Each dot is one of Michelson's measurements, and the grey line connects
the mean taken for each day. The same plot also shows simulated data
using a probability model. We'll get into statistics later in the
course; for now, let's focus on understanding what real and simulated
data tend to look like.

### **q4** Inspect the following plot with the `Real` Michelson data and `Simulated` data from a probability model. Document the similarities and differences between the data under *observe* below.

```{r q4-cf-real-simulated}
## Note: No need to edit this chunk!
## Calibrate simulated data
v_mean <-
  df_q2 %>%
  summarize(m = mean(VelocityVacuum)) %>%
  pull(m)
v_sd <-
  df_q2 %>%
  summarize(s = sd(VelocityVacuum)) %>%
  pull(s)

## Visualize
set.seed(101)
df_q2 %>%
  mutate(Simulated = rnorm(n(), mean = v_mean, sd = v_sd)) %>%
  rename(Real = VelocityVacuum) %>%
  pivot_longer(
    cols = c(Simulated, Real),
    names_to = "source",
    values_to = "velocity"
  ) %>%

  ggplot(aes(Date, velocity)) +
  geom_hline(
    yintercept = LIGHTSPEED_MICHELSON,
    linetype = "dotted"
  ) +
  geom_hline(
    yintercept = LIGHTSPEED_MICHELSON - LIGHTSPEED_PM,
    linetype = "dashed"
  ) +
  geom_hline(
    yintercept = LIGHTSPEED_MICHELSON + LIGHTSPEED_PM,
    linetype = "dashed"
  ) +

  geom_line(
    data = . %>%
      group_by(Date, source) %>%
      summarize(velocity_mean = mean(velocity)),
    mapping = aes(y = velocity_mean),
    color = "grey50"
  ) +
  geom_point(
    mapping = aes(y = velocity),
    size = 0.8
  ) +

  facet_grid(source~.) +
  theme_minimal() +
  labs(
    x = "Date of Measurement (1879)",
    y = "Velocity (in Vacuum)"
  )
```

**Observations**:

-   Similarities

    -   Both graphs have large portions of the trend line between the
        upper and lower bound of the data.

    -   Large spread of data points on any given day on either side of
        the average data line

-   Differences

    -   The peaks outside of the upper and lower bounds are on different
        dates.

    -   The actual points of data are different and the simulated looks
        like it has more variation

### **q5** You have access to a few other variables. Construct a **at least three** visualizations of `VelocityVacuum` against these other factors. Are there other patterns in the data that might help explain the difference between Michelson's estimate and `LIGHTSPEED_VACUUM`?

```{r q5-my-visuals}
df_q2 %>%
  ggplot(aes(VelocityVacuum, color = Distinctness)) +
  geom_density() +
  labs(title = "Density of Measured Velocity Values Per Distinctness", x = "Velocity of Light", y = "Density")

df_q2 %>% 
  ggplot(aes( x = Temp, y =VelocityVacuum)) +
  geom_smooth() +
  labs(title = "Trends of Velocity in Vacuum vs. Temperature", x = "Temperature", y = "Velocity of Light")
  
```

**Observations Part 1**:

-   Graph 1: Distinctness 3 is the highest quality data measurement. Yet
    its density is spread out rather than peaking towards one velocity.
    Distinctness 1 and 2 which are lower quality data measurements have
    distinct peaks where there is a clear trend along the data. This is
    the opposite of what I assumed would be true. I assumed the higher
    quality data would have less variation than the lower quality
    measurements. I could look into this as it could explain the
    difference between VelocityVacuum and Michelson's data.

-   Graph 2: Originally I made a smooth line graph of each distinctness.
    I ended up combining them into one trend line because there was too
    much variation in the data to make any accurate observations. This
    graph shows me that on the low range of data and high range (\<65
    and \>85) there is a more exponential relationship between
    temperature and the measured velocity in the vacuum. This is
    puzzling because in a vacuum we do not expect a change in speed of
    light due to temperature. In a median like air we might see a
    significantly small change due to the index of refraction. The trend
    on this graph is far more dramatic than that trend and comes from
    already untrustworthy data. We should not draw any claims about
    temperature having an effect on speed of light based on this data.
    Temperature does not affect the actual speed of light from my
    knowledge. From 65-85 degrees the velocity is tends to be consistent
    (in +- 15 from an average of about 299940). There is potential error
    in this smooth line graph as seen by the shaded regions and thus
    more analysis could be done.

```{r q5-day-of-week}
df_q5 <- 
  df_q2 %>%
  mutate(
    Day = wday(Date, label = TRUE)
  )

#plot density vs day of week
df_q5 %>%
  ggplot(aes(VelocityVacuum, color = Day)) +
  geom_density() +
  labs(title = "Density of Measured Velocity Vs. Day of The Week", x = "Velocity of Light", y = "Density")

df_q5 %>%
  ggplot(aes(x = Day, y = VelocityVacuum, fill = Day)) +
  geom_boxplot() +
  geom_hline(
    yintercept = LIGHTSPEED_VACUUM,
    linetype = "dotted"
  ) +
  labs(title = "Day of Week vs. Measured Velocity of Light Compared to Real Velocity of Light", x = "Day of The Week", y = "Velocity of Light")
```

**Observations Part 2**:

-   Graph 3: The trend density of velocity of the light is given for
    each day. Monday has the highest peak density yet has another bump.
    Wednesday is the most spread out over all the days meaning it tends
    to have the most variation in its data. Saturday, Thursday, and
    Tuesday follow similar trends with one peak at a velocity of
    roughly 299925. This shows that on those days they had the most
    similar data results. The data variation could potentially be due to
    different scientist taking the data on different days as most likely
    multiple people took data points. To understand the data trends per
    day we need to look at graph 4.
-   Graph 4: As stated in graph 3, Wednesday has the most variation as
    seen in its box plot. Monday and Thursday do not average around the
    same velocity but they have the least amount of variation. All box
    plots but Monday tend to have a mean data line around the same
    range. Monday has distinct separation from the other days. Thursday
    and Saturday have outlier data that is not take into account in the
    box plot and I wonder if that is due to instrumentation error and
    human error. The dotted line is the actual velocity of light. The
    only box + whisker plot that overlaps at all is Wednesday which is
    the day with the most variation. It shows that no matter the day all
    data tends to measure above the actual velocity of light. Day of the
    week could have a slight influence over the variation in data but
    does not explain why the data was so far from the actual velocity of
    light.

## Bibliography

-   [1] Michelson, [Experimental Determination of the Velocity of
    Light](https://play.google.com/books/reader?id=343nAAAAMAAJ&hl=en&pg=GBS.PA115)
    (1880) 
-   [2] Henrion and Fischhoff, [Assessing Uncertainty in Physical
    Constants](https://www.cmu.edu/epp/people/faculty/research/Fischoff-Henrion-Assessing%20uncertainty%20in%20physical%20constants.pdf)
    (1986) 
-   [3] BYU video about a [Fizeau-Foucault
    apparatus](https://www.youtube.com/watch?v=Ik5ORaaeaME), similar to
    what Michelson used.
