---
title: "Setup: RStudio Shortcuts"
author: Zachary del Rosario
date: 2020-05-07
output: github_document
time: 10
reading: 5
editor_options: 
  markdown: 
    wrap: 72
---

# Setup: RStudio Shortcuts

*Purpose*: Your ability to get stuff done is highly dependent on your
fluency with your tools. One aspect of fluency is knowing and *using*
keyboard shortcuts. In this exercise, we'll go over some of the most
important ones.

*Reading*: [Keyboard
shortcuts](https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts);
[Code Chunk Options](https://rmarkdown.rstudio.com/lesson-3.html)
*Note*: Use this reading to look up answers to the questions below.
Rather than *memorizing* this information, I recommend you download a
[cheatsheet](https://rstudio.com/wp-content/uploads/2016/01/rstudio-IDE-cheatsheet.pdf),
and either print it out or save it in a convenient place on your
computer. Get used to referencing your cheatsheets while doing data
science---practice makes perfect!

```{r setup, include=FALSE}
# knitr options
knitr::opts_chunk$set(echo = TRUE)
```

### **q1** What do the following keyboard shortcuts do?

-   Within the script editor or a chunk

    -   `Alt` + `-`

        Insert assignment operator

    -   `Shift` + `Cmd/Ctrl` + `M`

        Insert pipe operator

    -   `Cmd/Ctrl` + `Enter`

    Run current line/section

    -   `F1` (Note: on a Mac you need to press `fn` + `F1`)

    show help for function at cursor

    -   `Cmd/Ctrl` + `Shift` + `C`

    comment/uncomment current line / section

-   Within R Markdown

    -   `Cmd/Ctrl` + `Alt` + `I`

        Insert chunk

-   Within a chunk

    -   `Shift` + `Cmd/Ctrl` + `Enter`

        run current chunk

    -   `Ctrl` + `I` (`Cmd` + `I`)

        Reindent lines

Try this below!

```{r}
## Re-indent these lines
c(
  "foo",
  "bar",
  "goo",
  "gah"
)
```

### **q2** For a chunk, what header option do you use to

-   Run the code, don't display it, but show its results?

    echo = FALSE

-   Run the code, but don't display it or its results?

    include = FALSE

### **q3** How do stop the code in a chunk from running once it has started?

click the red square in the top right corner of the chunk

### **q4** How do you show the "Document Outline" in RStudio?

*Hint*: Try googling "rstudio document outline"

Use the keyboard shortcut `Ctrl + Shift + O` (Windows),
`Super + Shift + O` (Mac), or click the icon in the top-right of RStudio
(it’s between the Publish and Compass icons).

<!-- include-exit-ticket -->

# Exit Ticket

<!-- -------------------------------------------------- -->

Once you have completed this exercise, make sure to fill out the **exit
ticket survey**, [linked
here](https://docs.google.com/forms/d/e/1FAIpQLSeuq2LFIwWcm05e8-JU84A3irdEL7JkXhMq5Xtoalib36LFHw/viewform?usp=pp_url&entry.693978880=e-setup04-rstudio-shortcuts-assignment.Rmd).
