---
title: "Vis: Lines"
author: Zach del Rosario
date: 2020-06-04
output: github_document
time: 15
reading: 30
editor_options: 
  markdown: 
    wrap: 72
---

# Vis: Lines

*Purpose*: *Line plots* are a key tool for EDA. In contrast with a
scatterplot, a line plot assumes the data have a *function* relation.
This can create an issue if we try to plot data that do not satisfy our
assumptions. In this exercise, we'll practice some best-practices for
constructing line plots.

*Reading*: [Line plots](https://rstudio.cloud/learn/primers/3.6)
*Topics*: Welcome, Line graphs, Similar geoms (skip Maps) *Reading
Time*: \~30 minutes

```{r setup, include=FALSE}
# knitr options
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(tidyverse)
library(gapminder)

```

### **q1** The following graph doesn't work as its author intended. Based on what we learned in the reading, fix the following code.

```{r q1-task}
## TODO: Fix the following code
gapminder %>%
  filter(continent == "Asia") %>%
  ggplot(aes(year, lifeExp, color = country)) +
  geom_line()
```

### **q2** A line plot makes *a certain assumption* about the underlying data. What assumption is this? How does that assumption relate to the following graph? Put differently, why is the use of `geom_line` a bad idea for the following dataset?

```{r q2-vis}
## TODO: No need to edit; just answer the questions
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_line()

```

**Observations**: - What does a line plot assume about the data?  - How
does the `mpg` dataset violate this assumption? The line plot assumes
that there is one y value for every x value. The mpg dataset does not
have a function relationship between displ and hwy. Cars can have
identical displ with different hwy

### **q3** The following graph shows both the raw data and a smoothed version. Describe the trends that you can see in the different curves.

```{r q3-task}
## TODO: No need to edit; just interpret the graph
economics %>%
  ggplot(aes(date, unemploy)) +
  geom_line(aes(color = "Raw")) +
  geom_smooth(aes(color = "Smoothed"), se = FALSE) +
  scale_color_discrete(name = "Source")
```

**Observations**: - What can you tell based on the `Raw` data? - What
can you tell based on the `Smoothed` data?

Raw data shows a shorted cyclical patter while the smoothed version is a
longer-term trend occuring.

<!-- include-exit-ticket -->

# Exit Ticket

<!-- -------------------------------------------------- -->

Once you have completed this exercise, make sure to fill out the **exit
ticket survey**, [linked
here](https://docs.google.com/forms/d/e/1FAIpQLSeuq2LFIwWcm05e8-JU84A3irdEL7JkXhMq5Xtoalib36LFHw/viewform?usp=pp_url&entry.693978880=e-vis05-lines-assignment.Rmd).
