---
title: "Vis: Perceptual Basics"
author: Zach del Rosario
date: 2020-08-06
output: github_document
time: 20
reading: 40
---

# Vis: Perceptual Basics

*Purpose*: Creating a *presentation-quality* graph is an exercise in *communication*. In order to create graphs that other people can understand, we should know some stuff about *how humans see data*. Through the required "reading" (video) you'll learn about visual perception, then put these ideas to use criticizing some graphs. Later, you'll use these ideas to *improve* some graphs.

*Reading*: [How Humans See Data](https://www.youtube.com/watch?v=fSgEeI2Xpdc&list=PLluqivwOH1ouKkbM0c6x-g7DQnXF0UmC0&index=37&t=0s) (Video)

```{r setup}
library(tidyverse)
```

## Criticize these graphs!

<!-- -------------------------------------------------- -->

Using the ideas from the reading (video), state some issues with the following graphs. As a reminder, the *visual hierarchy* is:

1.  Position along a common scale
2.  Position on identical but nonaligned scales
3.  Length
4.  Angle; Slope (With slope not too close to 0, $\pi/2$, or $\pi$.)
5.  Area
6.  Volume; Density; Color saturation
7.  Color hue

### **q1** What are some issues with the following graph? *Don't just say* "it's bad"---use concepts from the required reading.

```{r q1-vis}
## NOTE: No need to edit; run and criticize
mpg %>%
  ggplot(aes(manufacturer, cty)) +
  geom_boxplot() +
  coord_flip()
```

**Observations**:

-   It is bland. It has no color so there is not much distinction between the values
-   The manufactures should be reranked in order of median cty value so that its easy to the eye to compare manufacturers

### **q2** What are some issues with the following graph? *Don't just say* "it's bad"---use concepts from the required reading.

```{r q2-vis}
## NOTE: No need to edit; run and criticize
as_tibble(mtcars) %>%
  mutate(model = rownames(mtcars)) %>%

  ggplot(aes(x = "", y = "", size = mpg)) +
  geom_point() +
  facet_wrap(~model)
```

-   Finding one manufacturer is not inherently easy
-   There is a x label and a y label which do not give me any information and should not be on the graph

### **q3** What are some issues with the following graph? *Don't just say* "it's bad"---use concepts from the required reading.

```{r q3-vis}
## NOTE: No need to edit; run and criticize
diamonds %>%
  ggplot(aes(clarity, fill = cut)) +
  geom_bar()
```

-   It is hard to compare the same cut over different clarity if not ideal. You can see the general color but its not an exact value
-   Count is not explanatory enough. Is it count of number of observations or cty? Are they summed up?
-   Bad to have to recode length

### **q4** What are some issues with the following graph? *Don't just say* "it's bad"---use concepts from the required reading.

```{r q4-vis}
## NOTE: No need to edit; run and criticize
diamonds %>%
  ggplot(aes(x = "", fill = cut)) +
  geom_bar() +
  coord_polar("y") +
  labs(x = "")
```

-   it should be in terms of percent. It shows the count up to the total number of observations
-   The color is find but text could go inside each of the things
-   The angle information is hard to understand

<!-- include-exit-ticket -->

# Exit Ticket

<!-- -------------------------------------------------- -->

Once you have completed this exercise, make sure to fill out the **exit ticket survey**, [linked here](https://docs.google.com/forms/d/e/1FAIpQLSeuq2LFIwWcm05e8-JU84A3irdEL7JkXhMq5Xtoalib36LFHw/viewform?usp=pp_url&entry.693978880=e-vis07-perception-assignment.Rmd).
